master thesis

### 2018-03-11 Paper - Latex/Paper - Skizze - v3.tex
Paper - Methods - simulate loss of information

      For the maximum parsimony analysis of the real data, all transition probabilities were equated.
        However, the used castor package \cite{Louca2017} offers the possibility to enter different 
        transition probabilities. \\

      In the simulation, we chose two beta distributions and a threshold that indicates the change 
        between states. \\
      Different thresholds with different beta distributions were simulated, with different distributions 
        of parasites and free-livings: 50\% P to 50\% FL, 40\% P to 60\% FL, 30\% P to 70\% FL and 
        20\% P to 80 \% FL (\todo{see results simulation, ref...}).

@@

    % \begin{wrapfigure}{r}{0.6\textwidth}
    %   \begin{center}
    %     \includegraphics[trim = 0mm 0mm 0mm 0mm, clip, width=0.58\textwidth]{Figures/Workflow-Simulation.png}
    %   \end{center}
    %   \caption{Overview of the Simulation}
    %   \label{fig:Simulation Workflow}
    % \end{wrapfigure}

@@

    %---------------------------------------------------------------------------------------------------
    %---------------------------------------------------------------------------------------------------
    \subsection{tag tree}
      \anmerkung{instead of 'tag tee' simulating states and transitions between them (Emanuel)} \\
      At this point we want one fully tagged tree, and one less tagged tree which looks like our real 
        data.\\
      Let's say the first specie (the root node) was free-living (start with a parasite without a host 
        makes no sence). For every transition from a node to his child, we take a random number from
        the father distribution. We decided that from the biological perspective a beta distribution
        reflects our transition probabilities best (see Figure 2.1 \todo{ref einfügen}). \\
      \anmerkung{why is that the case and why is that from a biological perspective? (Bernhard)} \\
      \anmerkungstext{I'd rather say that to ensure that the parameter of the binomial distribution is 
        restricted to the [0,1] interval, we model it... (Bernhard)} \\
      For example when our father node was free-living, then we take from the free-living beta
        distribution. Is the number under the threshold for beeing parasite, we get a change and tag
        the current node as parasite. Otherwise we tag it as free-living. \\
      With this procedure we traverse through the tree from the root to every leaf node. A part of
        this code you see here:
      \begin{lstlisting}[gobble=6]
        from numpy import random
        if father_tag == 0:
          # freeliving_distribution:
          new_random = random.beta(a=A_FL, b=B_FL)
        else:
          # parasite_distribution:
          new_random = random.beta(a=A_P, b=B_P)
        tag = 0         # $\rightarrow$ FL
        if new_random < percentage_parasites:
          tag = 1       # $\rightarrow$ P
      \end{lstlisting}
      \begin{figure}
        \caption{60\% Free-living - 40\% Parasites}
        \centering
          \includegraphics[width=0.5\textwidth]{Figures/40-60.png}
      \end{figure}
      \todo{Bessere Beschriftung, Plot neu erstellen! U.a. mit threshold} \\
      We save each tag with the associated node ID in a nodelist. \\
      \anmerkung{simulationg loss of information (1) states (2) topology/multifurcation (Emanuel)} \\
      The real tree has much less information, we have only information from some current species 
        (leaf nodes) and \todo{and probably negligible internal nodes}. \\
      To simulate our real tree we save for every node an empty placeholde except for some leaf nodes.
      There we save the states again. The amount of these unknown information is one parameter, which we
      got from our real tree. Or which we can change to \todo{...}
      \todo{Was hiervon gehört ind Methoden, was schon in Implementierung oder ganz woanders hin?}

    %---------------------------------------------------------------------------------------------------
    %---------------------------------------------------------------------------------------------------
    \subsection{multifurcate tree}
      \anmerkung{simulating loss of information (Emanuel)} \\
      Another parameter is the nature and strength of the multifunction of the tree, since we do not 
        have a binary tree in the real case. After several measurements and analyzes, which we explain 
        in \todo{section/chapter x}, \anmerkung{fit, justification (Emanuel)} we decided to use a $\frac{1}{x}$ distribution, where x is the depth 
        of a node. This means, how deaper we are, how less information we have. \\
      We traverse through the tree and pick a random number between 0 and 1. If random number is smaller 
        as our limit ($\frac{1}{x}$), than we forget the node and hang every child to the father node of 
        the current node. \\
      \anmerkung{poisson process $\rightarrow$ fit that distribution, include depth as a predictor, see if significant (Emanuel)}
      \begin{lstlisting}[gobble=6]
        from numpy import random
        from utilities import Helpers

        def get_non_binary_tree(subtree, nodelist):
          i = 0
          while i != len(subtree.clades):
            if subtree.clades[i].is_terminal():             # is leaf node?
              i += 1
            else:
              element = Helpers.find_element_in_nodelist(subtree.clades[i].name, nodelist)
              limit = get_limit(element[1])
              new_random = random.uniform()                 # choose if we want to delete ourselve
              if new_random < limit:                        # or new_random < 0.9:
                subtree.clades += subtree.clades[i].clades  # add children
                del subtree.clades[i]                       # delete internal node
              else:                                         # if we don't deleted ourselve go on with children
                get_non_binary_tree(subtree.clades[i], nodelist)  # otherwise the children are in the current clade array
                i += 1
          return

          def get_limit(depth):
            limit = 1 - 1 / ((depth + 3) / 4)
            if limit < 0.1:
              limit = 0.1
            return limit
      \end{lstlisting}
      % # current clades: clade_1 clade_2 ... clade_i-1 clade_i ... clade:_n
      % # add children of clade_i, delete clade_i
      % # new clades:clade_1 clade_2 ... clade_i-1 child_clade_1 ... child_clade_m ... clade:_n
      % # child_clade_1 is new clade i
      Wir lassen das Limit nicht beliebig klein werden, sondern beschränken es auf 0.1.

    %---------------------------------------------------------------------------------------------------
    %---------------------------------------------------------------------------------------------------
    \subsection{maximum parsimony algorithms}
      \subsubsection{Fitch maximum parsimony}
        Described from [COO98] + others ... - implemented for multifurcating trees

>>>>

      For the maximum parsimony analysis of the real data, all transition probabilities were equated.
        However, the used castor package \cite{Louca2017} offers the possibility to enter different 
        transition probabilities.
      \begin{wrapfigure}{r}{0.5\textwidth}
        \begin{center}
          \includegraphics[trim = 0mm 0mm 0mm 15mm, clip, width=0.48\textwidth]{Figures/40-60.png}
        \end{center}
        \caption{60\% Free-living - 40\% Parasites \\ red: parasites, blue: free-living, \\ the threshold is at 0.4}
        \label{fig:Beta distribution}
      \end{wrapfigure}

      In the simulation, we chose two beta distributions and a threshold that indicates the change 
        between states. \\
      Different thresholds with different beta distributions were simulated, with different distributions 
        of parasites and free-livings: 50\% P to 50\% FL, 40\% P to 60\% FL, 30\% P to 70\% FL and 
        20\% P to 80 \% FL (\todo{see results simulation, ref...}). Figure \ref{fig:Beta distribution} 
        shows one example of these. \\
      \todo{Plot neu erstellen: Achsenbeschriftung, threshold}

@@

@@

    %---------------------------------------------------------------------------------------------------
    %---------------------------------------------------------------------------------------------------
    \subsection{simulating states and transitions between them}
      The next step is to simulate the states of the nodes using the transitions. Again, we simulate 
        fully known states and then 'forget' everything but a few in the leaf nodes so that you can 
        later compare the reconstruction with the origin. \\
      The root node (ancestor of all subsequent species) is (of course) free-living. That means it will 
        start in the free-living beta distribution. Now traverse from the root to the leaf nodes, always 
        pulling out of the current distribution until you get above the threshold and the new node 
        changes state. \\
      To ensure that the parameter of the binomial distribution is restricted to the [0,1] interval, we 
        model it with a beta distribution as in Figure \ref{fig:Beta distribution}. \\

      After traversing through the tree, each state is saved in a nodelist associated with the node ID 
        which is the OTT from OTL. \\

      Here begins the simulation of the lost information. This is on the one hand the states and on the 
        other the topology of the tree. Some splits of nodes are unknown with which the tree is 
        multifurcated (explained in the following section \todo{pageref}). \\

      In the real tree, there is usually only information about species living today $\rightarrow$ leaf 
        nodes. And beyond only a small percentage of these. All information about the states of the 
        internal node and one leaf node is 'forgotten' and stored in another column to the node. \\
      Different percentages of forgetting the information were simulated, as you can read in the 
      \todo{section ... from the results}.

    %---------------------------------------------------------------------------------------------------
    %---------------------------------------------------------------------------------------------------
    \subsection{simulating loss of information of the tree topology}
      As previously explained, some divisions in the tree are not known, so the real tree is not binary.
        We simulate this multifurcation of the tree by \todo{...}.

      On the one hand, the strength of multifunctionation plays a role, as does the distribution. 
        Modeling was done as explained in \todo{sec x}.

      \todo{bisher ohne Begründung:!!!}
      we decided to use a $\frac{1}{x}$ distribution, where x is the depth of a node. This means, how 
      deaper we are, how less information we have. \\
      We traverse through the tree and pick a random number between 0 and 1. If random number is smaller 
        as our limit ($\frac{1}{x}$), than we forget the node and hang every child to the father node of 
        the current node. \\
      \anmerkung{poisson process $\rightarrow$ fit that distribution, include depth as a predictor, see if significant (Emanuel)}
      \begin{lstlisting}[gobble=6]
        def get_non_binary_tree(subtree, nodelist):
          i = 0
          while i != len(subtree.clades):
            if subtree.clades[i].is_terminal():             # is leaf node?
              i += 1
            else:
              element = Helpers.find_element_in_nodelist(subtree.clades[i].name, nodelist)
              limit = get_limit(element[1])
              new_random = random.uniform()                 # choose if we want to delete ourselve
              if new_random < limit:                        # or new_random < 0.9:
                subtree.clades += subtree.clades[i].clades  # add children
                del subtree.clades[i]                       # delete internal node
              else:                                         # if we don't deleted ourselve go on with children
                get_non_binary_tree(subtree.clades[i], nodelist)  # otherwise the children are in the current clade array
                i += 1
          return

          def get_limit(depth):
            limit = 1 - 1 / ((depth + 3) / 4)
            if limit < 0.1:
              limit = 0.1
            return limit
      \end{lstlisting}
      Wir lassen das Limit nicht beliebig klein werden, sondern beschränken es auf 0.1. \\

      With this loss of information the reconstruction can start.

    %---------------------------------------------------------------------------------------------------
    %---------------------------------------------------------------------------------------------------
    %---------------------------------------------------------------------------------------------------
    \section{maximum parsimony algorithms}

      %---------------------------------------------------------------------------------------------------
      %---------------------------------------------------------------------------------------------------
      \subsection{Fitch maximum parsimony}
        Described from [COO98] + others ... - implemented for multifurcating trees
